Viscous Erosion in Porous Media

### 2018-03-28 docs/methods/erosion_methods.tex
NM

These fixed-area simulations enable us to precisely characterize the changing geometry as it approaches a terminal state. In particular, we measure the aspect ratio of the body (the width to height) and the opening angle of the front/rear corners. In these tests, we keep the spatial resolution fixed, $N_\iin = 1024$, and show three different cases of the smoothing parameters to assess how the numerical regularization influences shape. In all three cases, we keep the two parameters equal $\eps = \sigma$ and set them to the values 20/1024, 10/1024, and 5/1024. Figure~\ref{fig:arangle}(a) shows the aspect ratio as it varies in time. The initially circular geometry has an aspect ratio of unity, and, as erosion carves a more slender form, the aspect ratio increases monotonically. The three simulations are all very close to one another, demonstrating convergence as $\eps, \sigma \to 0$. In the simulation with these least smoothing ($\eps = \sigma = 5/1024$), the aspect ratio tends to a value of 2.7 in long time.

Next, we aim to measure the opening angle of the front/rear corner. Since these simulated bodies are actually smooth, we must have a procedure to estimate the apparent angle that results from the high (but finite) curvature at the stagnation points. To this end, we fit the tangent angle $\theta(\alpha)$ with a 7th-degree polynomial over a region excluding the stagnation points ($\alpha = 0$ and 0.5). We then extrapolate the polynomial to the stagnation points, which gives an estimate of the angle there. Although this procedure is subject to some uncertainty (to be examined below), we have found it to give robust angle measurements. In Fig.~\ref{fig:arangle}(b), we show the opening angle on the eroding body for the same three cases of smoothing parameters. 
The method gives an angle measurement of $180^{\circ}$ for the initially circular geometry, thus providing some confirmation. As time moves forward, erosion carves a sharper corner and the angle decreases monotonically. Here, the effect of the smoothing parameters is more visible, but we still see convergence as $\eps, \sigma \to 0$. In the most accurate simulation (black), the opening angle converges to a value of $102^{\circ}$, which agrees precisely with the analytical prediction from Section \ref{LimitingShape}.

%As seen in Fig.~\ref{fig:arangle}a, the opening angle is initially $180^\circ$, corresponding to the initially circular geometry. Erosion causes the angle to form after a short time, and the angle becomes increasingly sharp as it converges to the value $102^\circ$ in the most resolved simulation. This coincides exactly with the valued predicted by our local solutions.

To examine in more detail the influence of the smoothing parameters, we show in Table

We now examine in detail the influence of the smoothing parameters and the uncertainty in the angle measurements...

@@

To examine the influence of the smoothing parameters, $\eps$ and $\sigma$, we show in Table~\ref{table:arangle} the measurements of the aspect ratio and the opening angle for different simulations. We fix the total number of discretization points at $N_\iin = 1024$ and sequentially reduce the smoothing with $\eps$ and $\sigma$ always equal. We choose the values $20/1024$, $10/1024$, and $5/1024$.  Loosely, this corresponds to smoothing over the nearest 20, 10, and 5 grid points respectively. As seen in the table, the aspect ratio converges to a values of $2.65$ and the opening angle converges to $102^{\circ}$. We also estimate the uncertainty in the angle measurements by varying the region of the polynomial fit and the degree of the fit. As shown in the table, as the smoothing parameters are reduced, the angle measurements become more precise.

>>>>

	These fixed-area simulations enable us to precisely characterize the changing geometry as it approaches a terminal state. In particular, we measure the aspect ratio of the body (the width to height) and the opening angle of the front/rear corners. In these tests, we keep the spatial resolution fixed, $N_\iin = 1024$, and show three different cases of the smoothing parameters to assess how the numerical regularization influences shape. In all three cases, we keep the two parameters equal $\eps = \sigma$ and set to the values 20/1024, 10/1024, and 5/1024 respectively. Figure~\ref{fig:arangle}(a) shows the aspect ratio as it varies in time. The initially circular geometry has an aspect ratio of unity, and, as erosion carves a more slender form, the aspect ratio increases monotonically. The three simulations are all very close to one another, demonstrating convergence as $\eps, \sigma \to 0$. In the simulation with these least smoothing ($\eps = \sigma = 5/1024$), the aspect ratio tends to a value of 2.7 in long time.

	Next, we aim to measure the opening angle of the front/rear corner. Since these simulated bodies are actually smooth on a fine scale, we must have a procedure to estimate the apparent angle that results from the high (but finite) curvature at the stagnation points. To this end, we fit the tangent angle $\theta(\alpha)$ with a 7th-degree polynomial over a region excluding the stagnation points ($\alpha = 0$ and 0.5). We then extrapolate the polynomial to the stagnation points, which gives an estimate of the angle there. Although this procedure is subject to some uncertainty (to be examined below), we have found it to give robust angle measurements. Figure \ref{fig:arangle}(b) shows the opening angle, $\beta$, as it varies in time for the three cases of smoothing parameters. For the initially circular geometry, our procedure yields the correct result of a $180^{\circ}$ opening angle, i.e.~a flat front face. As time proceeds, erosion carves a sharper corner and $\beta$ decreases monotonically. Here, the three curves are more easily distinguished than in Fig.~\ref{fig:arangle}(a), implying that the opening angle is more sensitive to the level of numerical smoothing. Nonetheless, the curves show convergence as $\eps, \sigma \to 0$. In the finest simulation ($\eps = \sigma = 5/1024$), $\beta$ converges to the value $102^{\circ}$, which agrees precisely with the analytical prediction from Section \ref{LimitingShape}.

	To more thoroughly examine the influence of the smoothing parameters, we report in Table~\ref{table:arangle} the long-time values of the aspect ratio and opening-angle measurements for the three cases. Loosely, with $N_\iin = 1024$ fixed, the values 20/1024, 10/1024, and 5/1024 correspond to smoothing over the nearest 20, 10, and 5 grid points respectively. As seen in the table, both the aspect ratio and the opening angle converge as $\eps, \sigma \to 0$ with roughly first order. To quantify the uncertainty involved in our angle measurement procedure, we show the estimated uncertainty, which is obtained by varying the degree of the fitting polynomial and the size of the excluded region around the stagnation point....

@@

%We fix the total number of discretization points at $N_\iin = 1024$ and sequentially reduce the smoothing with $\eps$ and $\sigma$ always equal. 
%Loosely, this corresponds to smoothing over the nearest 20, 10, and 5 grid points respectively. 
%As seen in the table, the aspect ratio converges to a values of $2.65$ and the opening angle converges to $102^{\circ}$. We also estimate the uncertainty in the angle measurements by varying the region of the polynomial fit and the degree of the fit. As shown in the table, as the smoothing parameters are reduced, the angle measurements become more precise.
